---
title: "Analyzing annual data"
author: "Abby Lewis"
date: "2022-09-21"
output: html_document
---

Load packages and lake data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(lubridate)
library(ggthemes)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
#remotes::install_github("ropensci/rnaturalearthhires")
library(rnaturalearthhires)
library(ggspatial)
library(ggpubr)
source("sen_slope_custom.R")
source("map trends and vals.R")
source("monthly_correlations.R")

with_temp <- read.csv("../Compiled data/All_data_annual AF.csv")
lat_long <- read.csv("https://pasta.lternet.edu/package/data/eml/edi/1530/1/fadd3eaa25b5fdd1fc4efba70e660579")

```


Load TP input data
```{r}
wi_lake_tp_conc <- readxl::read_excel("../External data/WI lakes TP.xlsx", sheet = 1) %>%
  pivot_longer(-sample_dt, names_to = "id", values_to = "TP_conc")
wi_lake_tp_flux <- readxl::read_excel("../External data/WI lakes TP.xlsx", sheet = 2) %>%
  pivot_longer(-sample_dt, names_to = "id", values_to = "TP_flux")
wi_lake_ids <- read.csv("../External data/WI_lake47.txt", sep = "\t")
wi_lake_areas <- read.csv("../External data/static_attributes.csv")

wi_tp_withID <- wi_lake_tp_conc %>%
  full_join(wi_lake_tp_flux) %>%
  left_join(wi_lake_ids, by = c("id")) %>%
  rename(hylak_id = basin_id) %>%
  left_join(wi_lake_areas %>% dplyr::select(basin, area_km2), by = c("hylak_id" = "basin"))  %>%
  left_join(lat_long) %>%
  filter(sample_dt > as.POSIXct("1981-01-01")) %>%
  mutate(TP_flux_kg_d = (TP_flux/1000000)*area_km2*1000000, #mg/m2/d --> g/d
         Volume_L = (SurfaceArea_ha*10000 * MeanDepth_m)*1000,
         TP_flux_ug_L_d = (TP_flux_kg_d*1000*1000*1000) / Volume_L)
```


Map summary stats for inputs
```{r}
summary_stats = wi_tp_withID %>%
  pivot_longer(c(TP_flux, TP_conc, TP_flux_kg_d, TP_flux_ug_L_d)) %>%
  group_by(LakeID, name) %>%
  summarize(mean = mean(value, na.rm = T),
            median = median(value, na.rm = T))

mean_map_flux_g_d = map_val(summary_stats %>% filter(name == "TP_flux_kg_d"), "TP flux mass\n(kg/d)")
mean_map_conc = map_val(summary_stats %>% filter(name == "TP_conc"), "TP conc\n(mg/L)")
mean_map_flux_ug_L_d = map_val(summary_stats %>% filter(name == "TP_flux_ug_L_d"), "TP flux vol\n(ug/L/d)")
mean_map_flux_mg_m2_d = map_val(summary_stats %>% filter(name == "TP_flux"), "TP flux area\n(mg/m2/d)")

summary_stats %>% 
  mutate(name = ifelse(name == "TP_flux", "TP_flux_mg_m2_d", name)) %>%
  ggplot(aes(x = mean)) +
  geom_histogram()+
  facet_wrap(~name, scales = "free")

jpeg("../Figures/TP_input_means.jpeg", width = 6, height = 4, units = "in", res = 300)
ggarrange(mean_map_flux_mg_m2_d, mean_map_conc, mean_map_flux_g_d, mean_map_flux_ug_L_d)
dev.off()
```


Calculate trends and map
Take home: inputs are decreasing
```{r}
wi_tp_trend <- wi_tp_withID %>%
  rename(Date = sample_dt) %>%
  mutate(Year = year(Date)) %>%
  group_by(Year, LakeID) %>%
  summarize(TP_flux = mean(TP_flux, na.rm = T),
            TP_conc = mean(TP_conc, na.rm = T),
            TP_flux_kg_d = mean(TP_flux_kg_d, na.rm = T),
            TP_flux_ug_L_d = mean(TP_flux_ug_L_d, na.rm = T)) %>%
  mutate(date = as.Date(paste0(Year, "-01-01")))

jpeg("../Figures/TP_input_patterns.jpeg", width = 6, height = 4, units = "in", res = 300)
wi_tp_trend %>%
  group_by(LakeID) %>%
  mutate(z_score = (TP_flux - mean(TP_flux, na.rm = T))/sd(TP_flux, na.rm = T)) %>%
  ggplot(aes(x = Year, y = z_score))+
  geom_line(aes(group = LakeID))+
  theme(legend.position = "none")+
  geom_smooth() +
  ylab("Z-score of TP input at each lake")+
  xlab("Year")+
  theme_bw()
dev.off()

jpeg("../Figures/TP_conc_patterns.jpeg", width = 6, height = 4, units = "in", res = 300)
with_temp %>%
  group_by(LakeID) %>%
  filter(StateOrProvince == "Wisconsin",
         Year >= 1990) %>%
  mutate(z_score = (TP_ugL_SURF - mean(TP_ugL_SURF, na.rm = T))/sd(TP_ugL_SURF, na.rm = T)) %>%
  ggplot(aes(x = Year, y = z_score))+
  geom_line(aes(group = LakeID))+
  theme(legend.position = "none")+
  geom_smooth() +
  ylab("Z-score of TP concentration at each lake")+
  xlab("Year")+
  theme_bw()
dev.off()

epi_tp_years <- with_temp %>%
  filter(!is.na(TP_ugL_EPI)) %>%
  group_by(LakeID) %>%
  summarize(min_year = min(Year),
            max_year = max(Year))

tp_input_trend_flux <- sen_slope_custom(wi_tp_trend, "TP_flux")
tp_input_trend_flux_kg_d <- sen_slope_custom(wi_tp_trend, "TP_flux_kg_d")
tp_input_trend_flux_ug_Ld <- sen_slope_custom(wi_tp_trend, "TP_flux_ug_L_d")
tp_input_trend_conc <- sen_slope_custom(wi_tp_trend, "TP_conc")

us_map_flux <- map_trend(tp_input_trend_flux, "TP flux area\n(mg/m2/d)")
us_map_conc <- map_trend(tp_input_trend_conc, "TP conc\n(mg/L)")
us_map_flux_kg_d <- map_trend(tp_input_trend_flux_kg_d, "TP flux mass\n(kg/d)")
us_map_flux_ug_L_d <- map_trend(tp_input_trend_flux_ug_Ld, "TP flux vol\n(µg/L/d)")

jpeg("../Figures/TP_input_trends.jpeg", width = 6, height = 4, units = "in", res = 300)
ggarrange(us_map_flux, us_map_conc, us_map_flux_kg_d, us_map_flux_ug_L_d)
dev.off()

#Make a flux dataset constrained to the years of in-lake data availability
input_limited_epi_tp <- wi_tp_trend %>%
  left_join(epi_tp_years) %>%
  filter(Year >= min_year & Year <= max_year)

tp_input_trend_flux_limited <- sen_slope_custom(input_limited_epi_tp, "TP_flux")
tp_input_trend_flux_kg_d_limited <- sen_slope_custom(input_limited_epi_tp, "TP_flux_kg_d")
tp_input_trend_flux_ug_Ld_limited <- sen_slope_custom(input_limited_epi_tp, "TP_flux_ug_L_d")
tp_input_trend_conc_limited <- sen_slope_custom(input_limited_epi_tp, "TP_conc")

input_limited_epi_tp %>%
  ggplot()+
  geom_density(aes(x = min_year)) +
  geom_density(aes(x = max_year))

#Make a flux dataset constrained to 2000 - 2020
input_limited_epi_tp_years <- wi_tp_trend %>%
  filter(Year >= 2000 & Year <= 2020)

tp_input_trend_flux_limited_years <- sen_slope_custom(input_limited_epi_tp_years, "TP_flux")
tp_input_trend_flux_kg_d_limited_years <- sen_slope_custom(input_limited_epi_tp_years, "TP_flux_kg_d")
tp_input_trend_flux_ug_Ld_limited_years <- sen_slope_custom(input_limited_epi_tp_years, "TP_flux_ug_L_d")
tp_input_trend_conc_limited_years <- sen_slope_custom(input_limited_epi_tp_years, "TP_conc")

summary_stats2 <- wi_tp_trend %>%
  left_join(epi_tp_years) %>%
  filter(Year >= min_year & Year <= max_year) %>%
  pivot_longer(c(TP_flux, TP_conc, TP_flux_kg_d, TP_flux_ug_L_d)) %>%
  mutate(name = paste0(name, "_limited")) %>%
  group_by(LakeID, name) %>%
  summarize(mean = mean(value, na.rm = T),
            median = median(value, na.rm = T))

summary_stats3 <- wi_tp_trend %>%
  left_join(epi_tp_years) %>%
  filter(Year >= 2000 & Year <= 2020) %>%
  pivot_longer(c(TP_flux, TP_conc, TP_flux_kg_d, TP_flux_ug_L_d)) %>%
  mutate(name = paste0(name, "_limited_years")) %>%
  group_by(LakeID, name) %>%
  summarize(mean = mean(value, na.rm = T),
            median = median(value, na.rm = T))

all_trends_pct <- tp_input_trend_flux %>% mutate(name = "TP_flux") %>%
  full_join(tp_input_trend_conc %>% mutate(name = "TP_conc")) %>%
  full_join(tp_input_trend_flux_kg_d %>% mutate(name = "TP_flux_kg_d")) %>%
  full_join(tp_input_trend_flux_ug_Ld %>% mutate(name = "TP_flux_ug_L_d")) %>%
  full_join(tp_input_trend_conc_limited %>% mutate(name = "TP_conc_limited")) %>%
  full_join(tp_input_trend_flux_limited %>% mutate(name = "TP_flux_limited")) %>%
  full_join(tp_input_trend_conc_limited_years %>% mutate(name = "TP_conc_limited_years")) %>%
  full_join(tp_input_trend_flux_limited_years %>% mutate(name = "TP_flux_limited_years")) %>%
  left_join(summary_stats %>%
              full_join(summary_stats2) %>%
              full_join(summary_stats3)) %>%
  mutate(trend = trend / mean * 100)

us_map_flux_pct <- map_trend(all_trends_pct %>% filter(name == "TP_flux"), "TP flux trend\n(%/decade)")
us_map_conc_pct <- map_trend(all_trends_pct %>% filter(name == "TP_conc"), "TP conc. trend\n(%/decade)")

jpeg("../Figures/TP_input_trends_pct.jpeg", width = 6, height = 4, units = "in", res = 300)
ggarrange(us_map_flux_pct, us_map_conc_pct)
dev.off()
```


Map in-lake trends and compare to inputs
```{r}
summer_hypo_tp_sen <- sen_slope_custom(with_temp %>% filter(!is.na(TP_ugL_HYPO)), "TP_ugL_HYPO")
summer_epi_tp_sen <- sen_slope_custom(with_temp %>% filter(!is.na(TP_ugL_EPI)), "TP_ugL_EPI")
surf_epi_tp_sen <- sen_slope_custom(with_temp %>% filter(!is.na(TP_ugL_SURF)), "TP_ugL_SURF")
hypo_tp_sen <- sen_slope_custom(with_temp %>% filter(!is.na(strat_TP_ugL_HYPO)), "strat_TP_ugL_HYPO")
epi_tp_sen <- sen_slope_custom(with_temp %>% filter(!is.na(strat_TP_ugL_EPI)), "strat_TP_ugL_EPI")
epi_chla_sen <- sen_slope_custom(with_temp %>% filter(!is.na(Chla_ugL_EPI)), "Chla_ugL_EPI")
hypo_do_sen <- sen_slope_custom(with_temp %>% 
                                  filter(!is.na(DO_mgL_HYPO)) %>%
                                  group_by(LakeID) %>%
                                  filter(mean(DO_mgL_HYPO) > 1) %>%
                                  ungroup(), 
                                "DO_mgL_HYPO")
af_sen <- sen_slope_custom(with_temp %>% filter(!is.na(AF)), "AF")

us_map_hypo_tp <- map_trend(summer_hypo_tp_sen %>% filter(!is.na(trend)), "Hypo. TP (µg/L/y)")
us_map_epi_tp <- map_trend(summer_epi_tp_sen %>% filter(!is.na(trend)), "Epi. TP (µg/L/y)")
us_map_hypo_tp_strat <- map_trend(hypo_tp_sen %>% filter(!is.na(trend)), "Hypo. TP (µg/L/y)")
us_map_epi_tp_strat <- map_trend(epi_tp_sen %>% filter(!is.na(trend)), "Epi. TP (µg/L/y)")
us_map_epi_tp_surf <- map_trend(surf_epi_tp_sen %>% filter(!is.na(trend)), "Surface TP (µg/L/y)")
us_map_epi_chla <- map_trend(epi_chla_sen %>% filter(!is.na(trend)), "Epi. chl-a (µg/L/y)")
us_map_af <- map_trend(af_sen %>% filter(!is.na(trend)), "AF (d)")
us_map_do <- map_trend(hypo_do_sen %>% filter(!is.na(trend)), "Hypo. DO (mg/L/d)")

tp_input_trend_flux_limited %>%
  mutate(var = "TP_flux_area") %>%
  dplyr::select(LakeID, trend, var) %>%
  full_join(epi_tp_sen %>% 
              mutate(var = "EPI_TP") %>%
              dplyr::select(LakeID, trend, var)) %>%
  pivot_wider(names_from = var, values_from = trend) %>%
  filter(!is.na(EPI_TP),
         !is.na(TP_flux_area)) %>%
  ggplot(aes(x = TP_flux_area, y = EPI_TP)) +
  geom_point()

# A quick note: I decided it wasn't relevant enough to include here, but we do see a correlation between epi. TP and epi. chl-a trends

wi_lakes_all_data <- with_temp %>%
  filter(LakeID %in% wi_tp_withID$LakeID) %>%
  mutate(date = ymd_hms(paste(Year,"-01-01 00:00:00")))

wi_tp_trend <- wi_tp_withID %>%
  rename(date = sample_dt)

all_lake_trends <- tp_input_trend_conc_limited %>%
  mutate(var = "TP_conc") %>%
  full_join(tp_input_trend_flux_limited %>% mutate(var = "TP_flux")) %>%
  full_join(epi_tp_sen %>% mutate(var = "strat_TP_ugL_EPI")) %>%
  full_join(surf_epi_tp_sen %>% mutate(var = "TP_ugL_SURF")) %>%
  full_join(hypo_tp_sen %>% mutate(var = "strat_TP_ugL_HYPO")) %>%
  full_join(hypo_do_sen %>% mutate(var = "DO_mgL_HYPO")) %>%
  full_join(af_sen %>% mutate(var = "AF")) %>%
  full_join(epi_chla_sen %>% mutate(var = "Chla_ugL_EPI"))

all_lake_trends_pct <- with_temp %>%
  pivot_longer(c(strat_TP_ugL_EPI, strat_TP_ugL_HYPO, TP_ugL_SURF, DO_mgL_HYPO, AF, Chla_ugL_EPI), names_to = "var") %>%
  group_by(LakeID, var) %>%
  summarize(mean = mean(value, na.rm = T),
            median = median(value, na.rm = T)) %>%
  left_join(all_lake_trends) %>%
  mutate(trend = trend / mean * 100)

jpeg("../Figures/All_trends.jpeg", width = 6, height = 6, units = "in", res = 300)
all_lake_trends %>%
  dplyr::select(LakeID, var, trend) %>%
  pivot_wider(names_from = var, values_from = trend) %>%
  pivot_longer(c(TP_flux, TP_conc), names_to = "input_var", values_to = "input_val") %>%
  pivot_longer(c(strat_TP_ugL_EPI, strat_TP_ugL_HYPO, Chla_ugL_EPI, DO_mgL_HYPO, AF), names_to = "lake_var", values_to = "lake_val") %>%
  filter(!is.na(input_val),
         !is.na(lake_val)) %>%
  ggplot(aes(x = input_val, y = lake_val))+
  geom_point()+
  ylab("Trend (in-lake variable)")+
  xlab("Trend (input)")+
  facet_grid(rows = vars(lake_var), cols = vars(input_var), scales = "free")
dev.off()
```

Make a figure for AGU to show that inputs are decreasing but concentrations are increasing
```{r}
#Calculate max and min to make scale bar consistent
all_lake_trends_pct_wi <- all_lake_trends_pct %>%
    left_join(lat_long) %>%
    filter(StateOrProvince == "Wisconsin")
comb_trends <- c(all_trends_pct$trend[all_trends_pct$name == "TP_flux_limited_years"],
                 all_lake_trends_pct_wi$trend[all_lake_trends_pct_wi$var == "strat_TP_ugL_EPI"])
abs_max <- max(abs(comb_trends), na.rm = T) * 10

us_map_flux_pct_agu <- map_trend(all_trends_pct %>% 
                                   filter(name == "TP_flux_limited_years",
                                          !is.na(trend)), 
                                 "")+
  scale_fill_gradientn(name = "Trend (% per decade)",
                       colours = c("blue","white","red"),
                       limits = c(-abs_max, abs_max),
                       n.breaks = 4) +
  ggtitle("TP input mass")

us_map_epi_tp_strat_pct <- map_trend(all_lake_trends_pct %>% 
                                   filter(var == "strat_TP_ugL_EPI",
                                          !is.na(trend)), 
                                 "")+
  scale_fill_gradientn(name = "Trend (% per decade)",
                       colours = c("blue","white","red"),
                       limits = c(-abs_max, abs_max),
                       n.breaks = 4) +
  ggtitle("In-lake surface-water TP")

jpeg("../Figures/TP_inlake_and_input_trends.jpeg", width = 6, height = 4, units = "in", res = 300)
ggarrange(us_map_flux_pct_agu, us_map_epi_tp_strat_pct, common.legend = TRUE, legend = "bottom")
dev.off()

all_lake_trends_pct %>% 
  filter(var == "strat_TP_ugL_EPI",
         !is.na(trend)) %>%
  full_join(all_trends_pct %>% 
              filter(name == "TP_flux_limited_years") %>%
              rename(var = name)) %>%
  ggplot(aes(x = trend, color = var)) +
  geom_density()

all_lake_trends_pct %>% 
  filter(var == "strat_TP_ugL_EPI") %>%
  full_join(all_trends_pct %>% 
              filter(name == "TP_flux_limited_years") %>%
              rename(var = name)) %>%
  select(LakeID, var, trend) %>%
  pivot_wider(names_from = var, values_from = trend) %>%
  filter(!is.na(TP_flux_limited_years), !is.na(strat_TP_ugL_EPI)) %>%
  ggplot(aes(x = TP_flux_limited_years, y = strat_TP_ugL_EPI)) +
  geom_point()

all_lake_trends_pct %>% 
  filter(var == "strat_TP_ugL_EPI") %>%
  full_join(all_trends_pct %>% 
              filter(name == "TP_flux_limited_years") %>%
              rename(var = name)) %>%
  select(LakeID, var, trend) %>%
  pivot_wider(names_from = var, values_from = trend) %>% 
  left_join(lat_long) %>%
  filter(StateOrProvince == "Wisconsin")

with_temp %>%
  filter(LakeID == "31")
```

Same figure, but using surface concentrations (i.e., including non-stratified lakes)
```{r}
#Calculate max and min to make scale bar consistent
all_lake_trends_pct_wi <- all_lake_trends_pct %>%
    left_join(lat_long) %>%
    filter(StateOrProvince == "Wisconsin")
comb_trends <- c(all_trends_pct$trend[all_trends_pct$name == "TP_flux"],
                 all_lake_trends_pct_wi$trend[all_lake_trends_pct_wi$var == "TP_ugL_SURF"])
abs_max <- max(abs(comb_trends), na.rm = T) * 10

us_map_flux_pct_agu <- map_trend(all_trends_pct %>% 
                                   filter(name == "TP_flux",
                                          !is.na(trend)), 
                                 "")+
  scale_fill_gradientn(name = "Trend (% per decade)",
                       colours = c("blue","white","red"),
                       limits = c(-abs_max, abs_max)) +
  ggtitle("TP input mass")

us_map_epi_tp_strat_pct <- map_trend(all_lake_trends_pct %>% 
                                   filter(var == "TP_ugL_SURF",
                                          !is.na(trend)), 
                                 "")+
  scale_fill_gradientn(name = "Trend (% per decade)",
                       colours = c("blue","white","red"),
                       limits = c(-abs_max, abs_max)) +
  ggtitle("In-lake surface-water TP")

jpeg("../Figures/TP_inlake_and_input_trends_unstrat.jpeg", width = 6, height = 4, units = "in", res = 300)
ggarrange(us_map_flux_pct_agu, us_map_epi_tp_strat_pct, common.legend = TRUE, legend = "bottom")
dev.off()

all_lake_trends_pct %>% 
  filter(var == "TP_ugL_SURF",
         !is.na(trend)) %>%
  full_join(all_trends_pct %>% 
              filter(name == "TP_flux") %>%
              rename(var = name)) %>%
  ggplot(aes(x = trend, color = var)) +
  geom_density()

all_lake_trends_pct %>% 
  filter(var == "TP_ugL_SURF") %>%
  full_join(all_trends_pct %>% 
              filter(name == "TP_flux") %>%
              rename(var = name)) %>%
  select(LakeID, var, trend) %>%
  pivot_wider(names_from = var, values_from = trend) %>%
  filter(!is.na(TP_flux), !is.na(TP_ugL_SURF)) %>%
  ggplot(aes(x = TP_flux, y = TP_ugL_SURF)) +
  geom_point()

all_lake_trends_pct %>%
  filter(var == "TP_ugL_SURF") %>%
  full_join(all_trends_pct %>% 
              filter(name == "TP_flux")%>%
              rename(var = name)) %>%
  select(LakeID, var, trend) %>%
  pivot_wider(names_from = var, values_from = trend) %>%
  filter(!is.na(TP_ugL_SURF), !is.na(TP_flux)) %>%
  left_join(lat_long) %>%
  ggplot(aes(y = TP_ugL_SURF, x = TP_flux, color = MaximumDepth_m > 6)) +
  geom_point()

all_lake_trends_pct %>%
  filter(var == "TP_ugL_SURF") %>%
  left_join(lat_long) %>%
  ggplot(aes(y = trend, x = MaximumDepth_m)) +
  geom_point()+
  scale_x_log10()
```


Plot correlation between mean in-lake and input concentrations (i.e., not trends)
Take home: no correlation
```{r}
#Calculate summary stats
in_lake_sum = with_temp %>%
  group_by(LakeID) %>%
  summarize(median_EPI = median(TP_ugL_EPI, na.rm = T), 
            median_HYPO = median(TP_ugL_HYPO, na.rm = T),
            mean_EPI = mean(TP_ugL_EPI, na.rm = T), 
            mean_HYPO = mean(TP_ugL_HYPO, na.rm = T))

summary_stats = wi_tp_withID %>%
  pivot_longer(c(TP_flux, TP_conc, TP_flux_kg_d, TP_flux_ug_L_d)) %>%
  group_by(LakeID, name) %>%
  summarize(mean = mean(value, na.rm = T),
            median = median(value, na.rm = T))

#EPI
summary_stats %>%
  left_join(in_lake_sum) %>%
  mutate(name = ifelse(name == "TP_conc", "TP conc\n(mg/L)", name),
         name = ifelse(name == "TP_flux", "TP flux area\n(mg/m2/d)", name),
         name = ifelse(name == "TP_flux_kg_d", "TP flux mass\n(kg/d)", name),
         name = ifelse(name == "TP_flux_ug_L_d", "TP flux vol\n(%/decade)", name)) %>%
  ggplot(aes(x = mean, y = mean_EPI)) +
  geom_point() +
  scale_x_log10()+
  facet_wrap(~name, scales = "free_x") +
  geom_smooth(method = "lm") + 
  stat_cor()

#HYPO
summary_stats %>%
  left_join(in_lake_sum) %>%
  mutate(name = ifelse(name == "TP_conc", "TP conc\n(mg/L)", name),
         name = ifelse(name == "TP_flux", "TP flux area\n(mg/m2/d)", name),
         name = ifelse(name == "TP_flux_kg_d", "TP flux mass\n(kg/d)", name),
         name = ifelse(name == "TP_flux_ug_L_d", "TP flux vol\n(%/decade)", name)) %>%
  ggplot(aes(x = mean, y = mean_HYPO)) +
  geom_point() +
  scale_x_log10()+
  facet_wrap(~name, scales = "free_x") +
  geom_smooth(method = "lm") + 
  stat_cor()
```


Monthly inputs correlation
Take home: on average no evidence for an effect of any one month of TP input on WI lake concentrations
```{r}
monthly_tp <- wi_tp_withID %>%
  mutate(mon = month(sample_dt),
         Year = year(sample_dt)) %>%
  group_by(LakeID, mon, Year) %>%
  summarise(TP_flux_sum = sum(TP_flux),
            TP_conc = mean(TP_flux)) %>%
  mutate(Year = ifelse(mon>=9,Year+1,Year))

wi_lakes_all_data <- with_temp %>%
  filter(LakeID %in% monthly_tp$LakeID) %>%
  full_join(monthly_tp)

monthly_correlations("TP_ugL_EPI", "TP_flux_sum", wi_lakes_all_data)
monthly_correlations("strat_TP_ugL_EPI", "TP_flux_sum", wi_lakes_all_data)
monthly_correlations("TP_ugL_HYPO", "TP_flux_sum", wi_lakes_all_data)
monthly_correlations("strat_TP_ugL_HYPO", "TP_flux_sum", wi_lakes_all_data)
monthly_correlations("strat_TP_ugL_HYPO", "TP_flux_sum", wi_lakes_all_data)
monthly_correlations("Chla_ugL_EPI", "TP_flux_sum", wi_lakes_all_data)

monthly_correlations("TP_ugL_EPI", "TP_conc", wi_lakes_all_data)
monthly_correlations("strat_TP_ugL_EPI", "TP_conc", wi_lakes_all_data)
monthly_correlations("TP_ugL_HYPO", "TP_conc", wi_lakes_all_data)
monthly_correlations("strat_TP_ugL_HYPO", "TP_conc", wi_lakes_all_data)
monthly_correlations("strat_TP_ugL_HYPO", "TP_conc", wi_lakes_all_data)
monthly_correlations("Chla_ugL_EPI", "TP_conc", wi_lakes_all_data)
```

Within-lake correlation plots
```{r}
annual_tp <- wi_tp_withID %>%
  mutate(mon = month(sample_dt),
         Year = year(sample_dt)) %>%
  mutate(Year = ifelse(mon>=9,Year+1,Year)) %>%
  group_by(LakeID, Year) %>%
  summarise(TP_flux = sum(TP_flux),
            TP_conc = mean(TP_conc))

wi_lakes_annual <- with_temp %>%
  filter(LakeID %in% annual_tp$LakeID) %>%
  full_join(annual_tp) %>%
  filter(!is.na(TP_ugL_EPI)) %>%
  group_by(LakeID)%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10)

wi_lakes_annual_surf <- with_temp %>%
  filter(LakeID %in% annual_tp$LakeID) %>%
  full_join(annual_tp) %>%
  filter(!is.na(TP_ugL_SURF)) %>%
  group_by(LakeID)%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10)

jpeg("../Figures/input correlation flux.jpeg", width = 8, height = 8, units = "in", res = 300)
wi_lakes_annual %>%
  group_by(LakeID)%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10) %>%
  ggplot(aes(x = TP_flux, y = TP_ugL_EPI, color = Year)) +
  scale_color_viridis_c()+
  geom_point()+
  scale_x_continuous(labels = ~ round(.x, 2))+
  facet_wrap(~LakeID, scales = "free")
dev.off()

jpeg("../Figures/input correlation flux surf.jpeg", width = 8, height = 8, units = "in", res = 300)
wi_lakes_annual_surf %>%
  group_by(LakeID)%>%
  filter(!is.na(TP_ugL_SURF), !is.na(TP_flux))%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10) %>%
  ggplot(aes(x = TP_flux, y = TP_ugL_SURF, color = Year)) +
  scale_color_viridis_c()+
  geom_point()+
  scale_x_continuous(labels = ~ round(.x, 2))+
  facet_wrap(~LakeName, scales = "free") +
  theme_bw()
dev.off()

jpeg("../Figures/input correlation flux surf select.jpeg", width = 5, height = 3, units = "in", res = 300)
wi_lakes_annual_surf %>%
  group_by(LakeID)%>%
  filter(!is.na(TP_ugL_SURF), !is.na(TP_flux))%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10,
         LakeName %in% c("Delavan", "Lac La Belle")) %>%
  ggplot(aes(x = TP_flux, y = TP_ugL_SURF, color = Year)) +
  scale_color_viridis_c()+
  geom_point()+
  scale_x_continuous(labels = ~ round(.x, 2))+
  facet_wrap(~LakeName, scales = "free") +
  theme_bw() +
  ylab("Surface TP concentration (µg/L)")+
  xlab("Annual TP input (kg/m2/d)")
dev.off()

jpeg("../Figures/input correlation conc.jpeg", width = 8, height = 8, units = "in", res = 300)
wi_lakes_annual %>%
  group_by(LakeID)%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10) %>%
  ggplot(aes(x = TP_conc, y = TP_ugL_EPI, color = Year)) +
  scale_color_viridis_c()+
  geom_point()+
  scale_x_continuous(labels = ~ round(.x, 2))+
  facet_wrap(~LakeID, scales = "free")
dev.off()

wi_lakes_annual <- with_temp %>%
  filter(LakeID %in% annual_tp$LakeID) %>%
  full_join(annual_tp) %>%
  filter(!is.na(Chla_ugL_EPI)) %>%
  group_by(LakeID)%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10)

jpeg("../Figures/input correlation flux chla.jpeg", width = 8, height = 8, units = "in", res = 300)
wi_lakes_annual %>%
  group_by(LakeID)%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10) %>%
  ggplot(aes(x = TP_flux, y = Chla_ugL_EPI, color = Year)) +
  scale_color_viridis_c()+
  geom_point()+
  scale_x_continuous(labels = ~ round(.x, 2))+
  facet_wrap(~LakeID, scales = "free")
dev.off()

jpeg("../Figures/input correlation conc chla.jpeg", width = 8, height = 8, units = "in", res = 300)
wi_lakes_annual %>%
  group_by(LakeID)%>%
  mutate(nyear = length(unique(Year)))%>%
  filter(nyear>=10) %>%
  ggplot(aes(x = TP_conc, y = Chla_ugL_EPI, color = Year)) +
  scale_color_viridis_c()+
  geom_point()+
  scale_x_continuous(labels = ~ round(.x, 2))+
  facet_wrap(~LakeID, scales = "free")
dev.off()
```

Regressions
```{r}
cummean.na <- function(x, na.rm = T) {
  n <- length(x)
  op <- rep(NA, n)
  for(i in 1:n) {op[i] <- mean(x[1:i], na.rm = !!na.rm)}
  rm(x, na.rm, n, i)
  return(op)
}

#Add relevant lags and seasonal calculations
with_lags <- with_temp%>%
  full_join(annual_tp) %>%
  unique()%>%
  group_by(LakeID)%>%
  arrange(LakeID,Year)%>%
  mutate(chla_lag = lag(Chla_ugL_EPI),
         epi_p_lag = lag(TP_ugL_EPI),
         strat_epi_p_lag = lag(strat_TP_ugL_EPI),
         hypo_p_lag = lag(TP_ugL_HYPO),
         epi_n_lag = lag(TN_ugL_EPI),
         hypo_n_lag = lag(TN_ugL_HYPO),
         summer_temp = (temp_july+temp_aug)/2,
         spring_temp = (temp_mar+temp_apr)/2,
         winter_temp = (temp_jan+temp_feb)/2,
         summer_precip = (precip_july+precip_aug)/2,
         spring_precip = (precip_mar +precip_apr)/2,
         winter_precip = (precip_jan +precip_feb)/2,
         anoxic = max(DO_mgL_HYPO,na.rm = T)<1,
         lag_is_last_year = ifelse(lag(Year) == (Year-1), T, F))%>%
  mutate(across(contains("_lag"), function(x) ifelse(lag_is_last_year, x, NA)))

# Save this as the dataset we will work with
dataset <- with_lags
row.names(dataset) <- paste(dataset$LakeID, dataset$Year)

data_log_nuts <- dataset%>%
  mutate(Chla_ugL_EPI=log(Chla_ugL_EPI),
         chla_lag = log(chla_lag),
         TP_ugL_EPI = log(TP_ugL_EPI),
         strat_TP_ugL_EPI = log(strat_TP_ugL_EPI),
         TP_ugL_HYPO = log(TP_ugL_HYPO),
         TN_ugL_EPI = log(TN_ugL_EPI),
         strat_TN_ugL_EPI = log(strat_TN_ugL_EPI),
         TN_ugL_HYPO = log(TN_ugL_HYPO),
         epi_p_lag = log(epi_p_lag),
         strat_epi_p_lag = log(strat_epi_p_lag),
         hypo_p_lag = log(hypo_p_lag),
         epi_n_lag = log(epi_n_lag),
         hypo_n_lag = log(hypo_n_lag), #produces NAs because of two EPI TN == 0
         TP_flux = log(TP_flux),
         TP_conc = log(TP_conc)) 

#Create datasets that are only oxic or only anoxic lakes
data_no_anoxic <- dataset%>%
  filter(anoxic==F)

data_always_oxic <- dataset %>%
  filter(!is.finite(max(AF, na.rm = T)) || max(AF, na.rm = T) == 0,
         !is.finite(min(DO_mgL_HYPO, na.rm = T)) || min(DO_mgL_HYPO, na.rm = T) > 1.7)

data_no_oxic <- dataset %>%
  filter(max(AF, na.rm = T) > 0)

source("lmer_functions.R")
library(lme4)
library(MuMIn)
library(ggridges)

### Epi. TP
responses <- c("strat_TP_ugL_EPI")
potential_drivers <- c("TP_ugL_HYPO",
                      "hypo_p_lag",
                      #"epi_p_lag",
                      "strat_buoyancy_freq_EPI",
                      "strat_TP_date_EPI",
                      "TP_flux"
                      )
aic_calculator_lmer(data_log_nuts,responses,potential_drivers, interaction = "+")
selected_drivers <- c("hypo_p_lag","TP_flux")
std_data <- standardize_data(data_log_nuts,responses,selected_drivers)
all_lakes_epiP <- mod_by_lake(data_log_nuts,responses,selected_drivers, interaction = "+")
mod_lmer <- lmer(strat_TP_ugL_EPI~hypo_p_lag+TP_flux+(1|LakeID), data = std_data)
vif(mod_lmer)
epi_p <- plot_effects_lmer(mod_lmer,"Epilimnetic TP (µg/L)", poster = F) 
epi_p_ridge <- plot_effects_by_lake_lmer_ridge(all_lakes_epiP,"Epilimnetic TP (µg/L)",mod_lmer, poster = F)

### TP with AF
responses <- c("TP_ugL_HYPO")
potential_drivers <- c("strat_TP_ugL_EPI",
                      "strat_buoyancy_freq_EPI",
                      "strat_Temp_C_HYPO",
                      "spring_precip",
                      "summer_precip",
                      "winter_precip",
                      "TP_date_HYPO",
                      "AF",
                      "TP_flux"
                      )
aic_calculator_lmer(data_log_nuts,responses,potential_drivers, interaction = "+")
selected_drivers <- c("TP_flux")
std_data <-standardize_data(data_log_nuts,responses,selected_drivers)
all_lakes_hypoP_af <- mod_by_lake(data_log_nuts,responses,selected_drivers, interaction = "+")
mod_lmer <- lmer(TP_ugL_HYPO~TP_flux+(1|LakeID), data = std_data)
vif(mod_lmer)
hypo_p_af <- plot_effects_lmer(mod_lmer,"Hypolimnetic TP (µg/L)", poster = F) 
hypo_p_ridge_af <- plot_effects_by_lake_lmer_ridge(all_lakes_hypoP_af,"Hypolimnetic TP (µg/L)",mod_lmer, poster = F)
lakes = data_log_nuts%>%
  ungroup()%>%
  dplyr::select(all_of(c(selected_drivers,responses,"LakeID","Year")))%>%
  na.omit()%>%
  group_by(LakeID)%>%
  filter(length(unique(Year))>=10)
lakes_using = unique(lakes$LakeID)
```
